---
title: "R as GIS: Interaction of Vector and Raster Datasets"
author: "AECN 396/896-002"
output:
  xaringan::moon_reader:
    # css: [default, metropolis, metropolis-fonts] 
    css: xaringan-themer.css 
    lib_dir: libs
    nature:
      ratio: 12:8
      highlightStyle: github
      highlightLines: true
      highlightSpans: true
      countIncrementalSlides: false
---

```{r, child = '../setup.Rmd'}
```

```{r, include = F, cache = F}
library(sf)
library(stars)
library(terra)
library(patchwork)
library(data.table)
library(mapview)
library(tidyverse)
library(raster)
library(here)
library(tmap)
library(exactextractr)

library(knitr)
opts_knit$set(error = TRUE, root.dir = here("Chapter-8-SpatialData"))
options(knitr.duplicate.label = "allow")
```

```{r echo = F, eval = F}
setwd("/Users/tmieno2/Dropbox/TeachingUNL/Data-Science-with-R/Chapter-8-SpatialData")
```
# Before you start

## Learning objectives

Learn the spatial interactions of a vector and raster dataset. Specifically,

.content-box-green[**Objective 1**]

Crop (spatially subset) a raster dataset based on the geographic extent of a vector dataset. 

.content-box-green[**Objective 2**]

Extract values from raster data for points and polygons. 

To be precise, here is what we mean by raster data extraction and what it does for points and polygons data:

+ **Points**: For each of the points, find which raster cell it is located within, and assign the value of the cell to the point.  
 
+ **Polygons**: For each of the polygons, identify all the raster cells that intersect with the polygon, and assign a vector of the cell values to the polygon

---
class: middle

# Before you start

## Table of contents

1. [Cropping raster datasets](#crop)
2. [Extract values from raster datasets](#extract)
3. [Basic speed comparison](#speed)

<br>

<span style="color:red"> Tips: </span>hitting letter "o" key will give you a panel view of the slides

---
class: middle

# Prerequisites 

.content-box-green[**Prerequisites**]

+ What Geographic Coordinate System (GCS), Coordinate Reference System (CRS), and projection are ([this](https://annakrystalli.me/intro-r-gis/gis.html) is a good resource)

+ Distinctions between vector and raster data ([this](https://gis.stackexchange.com/questions/57142/what-is-the-difference-between-vector-and-raster-data-models) is a simple summary of the difference)

---
class: middle

# Preparation 

+ Import treatment block shape files:

```{r results = "hide"}
treatment_blocks <- 
  st_read(dsn = "./Data", layer = "ZP_TRT_Blocks_Buff") %>%
  #--- transform the CRS so that it is consistent with the raster data ---#
  st_transform(4326)
```

+ Import yield point data

```{r results = "hide"}
yield <-
  st_read(dsn = "./Data", layer = "Stevens_Centuro_YLD_shp") %>%
  # dplyr::select(Yld_Vol_Dr) %>%
  # rename(dry_yield = Yld_Vol_Dr) %>%
  #--- create fake yield data ---#
  mutate(yield = runif(n()) * 240) %>%
  #--- transform the CRS so that it is consistent with the raster data ---#
  st_transform(4326)
```


+ Import drone imagery data (GeoTiff) file.

```{r , cache = FALSE}
ndre <- raster::raster("./Data/NDRE.tif")
ndre_sr <- terra::rast("./Data/NDRE.tif")
```

---
class: middle

.left4[
<br>

.content-box-green[**Note**]

+ gray circles: yield data points
+ red rectangles: treatment blocks
]

.right6[
```{r echo = F, out.width = "100%"}
tm_shape(ndre) +
  tm_raster(title = "NDRE") +
  tm_shape(yield) +
  tm_symbols(size = 0.02) +
  tm_shape(treatment_blocks) +
  tm_borders(col = "red") +
  tm_layout(legend.outside = TRUE)
```
]

---
class: inverse, center, middle
name: data-structure

# Cropping raster data 

<html><div style='float:left'></div><hr color='#EB811B' size=1px width=1000px></html>

---
class: middle

# Cropping raster data

+ Sometimes, it is convenient to crop a raster layer to the specific area of interest so that you do not have to carry around unnecessary parts of the raster layer. 

+ Moreover, it takes less time to extract values from a raster layer when the size of the raster layer is smaller. 

---
class: middle

.content-box-green[**How**]

You can crop a raster layer by using `raster::crop()` or `terra::crop()`.

```{r eval = F}
#--- raster ---#
raster::crop(Raster*, sf)

#--- terra ---#
terra::crop(SpatRaster, sf)
```

The resulting `Raster*` object is the original cropped to the bounding box of the `sf` object.

---
class: middle

.content-box-green[**Goal**]

Crop `ndre` to the bounding box of `treatment_blocks`.

.content-box-green[**Code**]

```{r cache = F}
#--- raster ---#
ndre_cropped <- raster::crop(ndre, treatment_blocks)

#--- terra ---#
ndre_cropped_sr <- raster::crop(ndre_sr, treatment_blocks)
```

---
class: middle

```{r echo = F, out.width = "80%"}
tm_shape(ndre_cropped) +
  tm_raster(title = "NDRE") +
  tm_shape(treatment_blocks) +
  tm_polygons(alpha = 0, size = 0.1) +
  tm_layout(legend.outside = TRUE)
```

---
class: inverse, center, middle
name: data-structure

# Extracting Values from Raster Layers for Vector Data

<html><div style='float:left'></div><hr color='#EB811B' size=1px width=1000px></html>

---
class: middle

# To Points

.content-box-green[**What**]

For each of the points, find which raster cell it is located within, and assign the value of the cell to the point.  

---
class: middle

```{r echo = F}
set.seed(378533)

#--- create polygons ---#
polygon <- st_polygon(list(
  rbind(c(0, 0), c(8, 0), c(8, 8), c(0, 8), c(0, 0))
))

raster_like_cells <- st_make_grid(polygon, n = c(8, 8)) %>%
  st_as_sf() %>%
  mutate(value = sample(1:64, 64))

stars_cells <- st_rasterize(raster_like_cells, nx = 8, ny = 8)

cell_centroids <- st_centroid(raster_like_cells) %>%
  st_as_sf()
```

.left4[
+ The numbers inside the cells are the values that the cells hold. 

+ After the extraction, 
	* Point 1 will be assigned $50$
	* Point 2 will be assigned $4$
	* Point 3 will be assigned $54$.
]

.right6[
```{r points-extact-viz, fig.cap = "Visual illustration of raster data extraction for points data", echo = F, out.width = "100%"}
#--------------------------
# Create points for which values are extracted
#--------------------------
#--- points ---#
point_1 <- st_point(c(2.4, 2.2))
point_2 <- st_point(c(6.7, 1.8))
point_3 <- st_point(c(4.2, 7.1))

#--- combine the points to make a single  sf of points ---#
points <- list(point_1, point_2, point_3) %>%
  st_sfc() %>%
  st_as_sf() %>%
  mutate(point_name = c("Point 1", "Point 2", "Point 3"))

#--------------------------
# Create maps
#--------------------------
ggplot() +
  geom_stars(data = stars_cells, alpha = 0.5) +
  scale_fill_distiller(name = "Value", palette = "Spectral") +
  geom_sf_text(data = raster_like_cells, aes(label = value)) +
  geom_sf(data = points, aes(shape = point_name), size = 2) +
  scale_shape(name = "Points") +
  theme_void()
```
]

---
class: middle

.content-box-green[**How**]

You can use `raster::extract()` or `terra::extract()` with the following syntax: 

```{r eval = F}
raster::extract(Raster^*, sf of points)
terra::extract(SpatRaster, sf of points)
```

---
class: middle


.content-box-green[**Goal**]

Extract NDRE values for yield points data:

+ match each of the yield points with the cell in which it resides in
+ assign the cell (NDRE) value to the point

---
class: middle

.content-box-green[**Implementation**]

```{r }
#--- raster ---#
ndre_extracted_r <- raster::extract(ndre, yield)

#--- terra ---#
ndre_extracted_t <- terra::extract(ndre_sr, yield)
```

.content-box-red[**Aside**]

On my computer, `terra::extract()` takes about 0.5 seconds while `raster::extract()` takes about 2.5 seconds.

.content-box-green[**Next step**]

Just extracting the raster values to the points is not where we stop. 

We need to merge the extracted values back to the points data so that we can use them for further analysis.

---
class: middle

Notice that the extracted data are in different format for the `raster` and `terra` options.

The `raster` way gives you a simple vector:    

```{r}
#--- raster ---#
head(ndre_extracted_r)
class(ndre_extracted_r)
```

The `terra` way gives you a `data.frame`

```{r}
#--- terra ---#
head(ndre_extracted_t)
class(ndre_extracted_t)
```

---
class: middle

# `raster` way

.content-box-green[**Merging the extracted values to the points data**]

The $n$th value in `ndre_sr_extracted_r` is for the $n$th point in `yield`.

So, you can simply do this:

```{r }
yield <- mutate(yield, NDRE = ndre_extracted_r)
```

We see only `NA` values for NDRE. No worries, the job was successful.

```{r }
filter(yield, !is.na(NDRE))
```

---
class: middle

# `terra` way

.content-box-green[**Merging the extracted values to the points data**]

The $n$th row in `ndre_sr_extracted_t` is for the $n$th point in `yield`.

So, you can simply do this:

```{r }
yield <- mutate(yield, NDRE = ndre_extracted_t$NDRE)
```

---
class: middle

# Multiple layers

Create a two-layer raster objects (two of the same layers here):

```{r, cache = FALSE}
ndre_stack <- raster::stack(c("./Data/NDRE.tif", "./Data/NDRE.tif"))
ndre_sr <- terra::rast(c("./Data/NDRE.tif", "./Data/NDRE.tif"))
```

Now, extract: 

```{r }
#--- raster ---#
ndre_extracted_r <- raster::extract(ndre_stack, yield)
head(ndre_extracted_r)
```

```{r}
#--- terra ---#
ndre_extracted_t <- terra::extract(ndre_sr, yield)
head(ndre_extracted_t)
```


---
class: middle

# To Polygons


.content-box-green[**What**]

For each of the polygons, identify all the raster cells that intersect with the polygon, and assign a vector of the cell values to the polygon. 

---
class: middle

.left4[

+ Find all the raster cells each of the polygons "intersect with"

+ Assign the value of all the intersecting cells to the polygon (n-to-1). 
]

.right6[
```{r polygons-extact-viz, fig.cap = "Visual illustration of raster data extraction for polygons data", echo = F, out.width = "100%"}
#--------------------------
# Create a polygon for which values are extracted
#--------------------------
polygon_extract <- st_polygon(list(
  rbind(c(1.5, 2), c(6, 2.3), c(7, 6.5), c(2, 5), c(1.5, 2))
))

polygons_extract_viz <- ggplot() +
  geom_stars(data = stars_cells, alpha = 0.5) +
  scale_fill_distiller(name = "Value", palette = "Spectral") +
  geom_sf(data = polygon_extract, fill = NA) +
  geom_sf(data = cell_centroids, color = "black", size = 0.8) +
  geom_sf_text(data = raster_like_cells, aes(label = value), nudge_x = -0.25, nudge_y = 0.25) +
  theme_void()

polygons_extract_viz
```
]

---
class: middle

.content-box-green[**How**]

We can use the `exact_extract()` function from the `exactextractr` package:

```{r eval = F}
library(exactextractr)
```

```{r eval = FALSE}
exact_extract(raster, polygons sf)
```

.content-box-green[**Why exact_extract()?**]

It is typically faster than `terra::extract()`.

---
class: middle

.content-box-green[**Goal**]

Extract NDRE values for the treatment block polygons.

.content-box-green[**Implementation**]

```{r results = "hide"}
ndre_extracted <- exact_extract(ndre, treatment_blocks)
```

---
class: middle

The object returned by `exact_extract()` is a list:

```{r }
#--- check the class ---#
class(ndre_extracted)
```

Since one list element is created for each of the polygons, the number of elements in the list is the same as the number of the polygons.

```{r }
#--- number of elements of the list ---#
length(ndre_extracted)

#--- number of the polygons ---#
nrow(treatment_blocks)
```

---
class: middle

Each element of the list is a `data.frame`, where extracted raster values (`value` column) and coverage fraction (`coverage_fraction` column) are stored.  

```{r }
head(ndre_extracted[[1]])
```

.content-box-green[**Key**]

$n$th element of the list is for the $n$th polygon in `treatment_blocks`.

---
class: middle

.content-box-green[**after-extraction processing**]

It is almost always the case that we do not need all of those raster values for further analysis. 

Rather, what we often need is the summary of the values by polygon (e.g., average).

To achieve this, you can use `bind_rows()` with `.id` option and then utilize `dplyr` operations.

```{r }
ndre_mean <- ndre_extracted %>%
  bind_rows(.id = "id") %>%
  #--- convert from character to numeric  ---#
  mutate(id = as.numeric(id)) %>%
  group_by(id) %>%
  summarize(NDRE = mean(value))
```

```{r echo = F}
head(ndre_mean)
```

In this dataset, the observation with `id == n` corresponds to $n$th polygon in `treatment_blocks`.

---
class: middle

.content-box-green[**Alternatively**]

You can use `fun` option inside `exact_extract()` to get a summary of raster values without you having to do it.

```{r results = "hide"}
ndre_extracted <- exact_extract(ndre, treatment_blocks, fun = "mean")
```

```{r }
head(ndre_extracted)
```

--

.content-box-green[**Note**]

Run `?exact_extract` to see other function options. 

---
class: middle


.content-box-green[**From a RasterStack**]

You can extract raster values from stacked raster layers by a single call of `exact_extract()` 

--

Let's create a stack:

```{r }
ndre_stack <- stack(ndre, ndre, ndre)
```

--

Extract from the stack:

```{r results = "hide"}
ndre_stack_extracted <- exact_extract(ndre_stack, treatment_blocks)
```

--

Let's take a look at the first element of the returned list.

```{r }
ndre_stack_extracted[[1]] %>% head()
```

Instead of having two columns, we now have three columns of extracted raster values and one column for `coverage_fraction`.

$n$th column is for raster values extracted from the $n$th layer.

---
class: inverse, center, middle
name: data-structure

# Appendix

<html><div style='float:left'></div><hr color='#EB811B' size=1px width=1000px></html>

---
class: middle

# `terra` package

.content-box-green[**Aside**]

The `terra` package (successor of the `raster` package) provides a faster extraction method with `terra::extract()`.

The `terra` package is faster than the `raster` package for all the raster operations. However, it is still in development and rather unstable at the moment.

Good news is that once you know how the `raster` package works, your transition to the `terra` package should be rather smooth.

.content-box-green[**Extracting for points**]

```{r }
#--- convert `RasterLayer` to `SpatRaster` ---#
ndre_sr <- rast(ndre)
```

```{r eval = F}
ndre_extracted <- terra::extract(ndre_sr, `st_coordinates`(yield))
```

```{r echo = F}
ndre_extracted <- terra::extract(ndre_sr, st_coordinates(yield))
```

---
class: inverse, center, middle
name: data-structure

# Appendix (terra)

<html><div style='float:left'></div><hr color='#EB811B' size=1px width=1000px></html>

---
class: middle

.content-box-green[**How**]

We can use the `extract()` function from the `raster` or `terra` package:

```{r eval = FALSE}
raster::extract(raster, polygons sf)
terra::extract(raster, polygons sf)
```

.content-box-green[**Note**]

Often times, `terra::extract()` is significantly faster than `raster::extract()`. It is more so when the underlying raster data is larger or there are more polygons.

---
class: middle

.content-box-green[**Goal**]

Extract NDRE values for the treatment block polygons.

.content-box-green[**Implementation**]

```{r results = "hide", cache = FALSE}
ndre_extracted <- terra::extract(ndre_sr, treatment_blocks)
```

---
class: middle

The object returned by `exact_extract()` is a list:

```{r }
#--- check the class ---#
class(ndre_extracted)
```

Notice that the observations with the same ID now have more than one values from the raster layers because you are extracting to polygons.

```{r}
head(ndre_extracted)
```
