---
title: "Course Logistics"
output:
  xaringan::moon_reader:
    # css: [default, metropolis, metropolis-fonts] 
    css: ../xaringan-themer.css 
    self_contained: true
    nature:
      ratio: 16:9
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---
class: middle

```{r, child = "setup.Rmd"}
```

```{r knitr-setup, include = FALSE, cache = F}
library(knitr)
library(here)
opts_chunk$set(
  echo = F,
  root.dir = here()
)
```

```{r prep, include = FALSE, cache = F}
library(data.table)
library(magick)
library(fixest)
library(officer)
library(flextable)
library(dplyr)
library(ggplot2)
```

# Meeting time

+ Tuesday: 8:50 am ~ 10:20 am
+ Friday: 8:50 am ~ 10:20 am

---
class: middle

# Some simple rules during a lecture

+ Keep your camera on
+ You do not ask me if you can go to a bathroom or not. Feel free to just leave
+ You are free to drink (no alcohol though), but not eat

---
class: middle

# Grading

+ Assignment 1: 40 points 
+ Assignment 2: 40 points
+ Final project: 20 points

---
class: middle

# Random thing

+ Any data sets that you want me to use for demonstrations?

---
class: middle

# Self-introduction


