---
title: "Data visualization with `ggplot2`"
author: "AECN 396/896-002"
output:
  xaringan::moon_reader:
    # css: [default, metropolis, metropolis-fonts] 
    css: xaringan-themer.css 
    nature:
      ratio: 16:10
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r, child = './../setup.Rmd'}
```

```{r, include = F, cache = F}
library(nycflights13)
library(extrafont)
library(data.table)
library(patchwork)
library(flipbookr)
```

# Before you start

## Learning objectives

The objectives of this chapter is to learn how to use the `ggplot2` package to create figures for effective communication 

## Table of contents

1. [`ggplot2` basics](#ggplot2-basics)
2. [Different types of figures](#dif-geoms)
3. [Placing more information in one figure](#more-info)
4. [Faceted figures](#faceting)
5. [Other supplementary `geom_*()`](#other-geoms)


<br>

<span style="color:red"> Tips: </span>hitting letter "o" key will give you a panel view of the slides


---

# `ggplot2` package

.left-full[

Install the package if you have not.

```{r ggplot2-install, eval = FALSE}
install.packages("ggplot2")
```

When you load the `tidyverse` package, it automatically load it. 

```{r tidyverse, message = TRUE}
#--- load ggplot2 along with others in the tidyverse package ---#
library(tidyverse)

#--- or ---#
library(ggplot2) #<<
```

]

---

# The datasets we use

```{r eval = F, echo = F}
setwd(here("Chapter-4-DataVisualization"))
```

.panelset[ 

.panel[.panel-name[Instruction]
Go [here](https://www.dropbox.com/sh/63rlp4ydmyjm1ui/AACYSeN0f_WAgKPQKzgpGVe0a?dl=0) and download **county_yield.rds** and then read the file onto R:
  ]

.panel[.panel-name[R Code]
```{r dataset, results = "hide", cache = F}
county_yield <-
  readRDS("county_yield.rds") %>%
  dplyr::select(
    soy_yield, corn_yield, year, county_code, state_name, d0_5_9, d1_5_9, d2_5_9, d3_5_9, d4_5_9
  ) %>%
  filter(state_name %in% c("Nebraska", "Kansas", "Colorado"))
```
  ]

.panel[.panel-name[Output]

```{r echo = F}
county_yield
```

  ]

.panel[.panel-name[Variable Definitions]
+ `soy_yield`: soybean yield (bu/acre)
+ `corn_yield`: corn yield (bu/acre)
+ `d0_5_9`: ratio of weeks under drought severity of 0 from May to September
+ `d1_5_9`: ~ drought severity of 1 from May to September   
+ `d2_5_9`: ~ drought severity of 2 from May to September   
+ `d3_5_9`: ~ drought severity of 3 from May to September   
+ `d4_5_9`: ~ drought severity of 4 from May to September   
  ]
]

<!-- 
#=========================================
# ggplot2 Basics
#=========================================
-->
  
---

class: inverse, center, middle
name: ggplot2-basics

# `ggplot2` basics

<html><div style='float:left'></div><hr color='#EB811B' size=1px width=1000px></html>

---
```{r taste, cache = F, include = F}
county_yield %>%
  ggplot(data = .) +
  aes(x = factor(year)) +
  aes(y = corn_yield) +
  geom_boxplot() +
  aes(fill = state_name) +
  facet_grid(state_name ~ .) +
  xlab("Year") +
  ylab("Corn Yield (bu/acre)") +
  ylim(c(100, 200)) +
  scale_fill_viridis_d(
    name = "State",
    guide = guide_legend(
      title.position = "left"
    )
  ) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, size = 6)) +
  theme(axis.text.y = element_text(size = 6)) +
  theme(legend.position = "bottom") +
  theme(
    legend.title = element_text(size = 6),
    legend.text = element_text(size = 6)
  ) +
  labs(title = "Corn Yield (bu/acre) by State") +
  labs(caption = "Design: Taro Mieno") +
  labs(subtitle = "Data Source: USDA-NASS")
```

`r chunk_reveal(
  chunk_name = "taste", 
  title = "# Taste of ggplot2",
  #break_type = "user", 
  display_type = "both",
  width = c(50, 50)
  )
`

---

# Basics

.panelset[ 

.panel[.panel-name[Step 1]

.left-full[

The very first job you need to do to create a figure using the `ggplot2` package is to let R know the dataset you are trying to visualize, which can be done using `ggplot()` as follows:

```{r g_fig_base}
g_fig <- ggplot(data = county_yield)
```

When you create a figure using the `ggplot2` package, `ggplot()` is always the function you call first.

]

  ]

.panel[.panel-name[g_fig]

.left-code[
Let's now see what is inside `g_fig`:

```{r g-fig, eval = F}  
g_fig
```

Well, it's blank. Obviously, `g_fig` still does not have enough information to create any kind of figures. You have not told R anything specific about how you would like to use the information in the dataset.
]

.right-plot[
```{r ref.label = "g-fig", echo = F, out.width = "90%"}  

```
]

  ]

.panel[.panel-name[Step 2]

.left-full[

So, the next thing you need to do is tell `g_fig` what type of figure you want by `geom_*()` functions. 

For example, we use `geom_point()` to create a scatter plot. 

To create a scatter plot, R needs to know which variables should be on the y-axis and x-axis. 

These information can be passed to `g_fig` by the following code:

```{r g_fig_point}
g_fig_scatter <- g_fig + geom_point(aes(x = d3_5_9, y = corn_yield))
```

Here, 

+ `geom_point()` was added to `g_fig` to declare that you want a scatter plot 
+ `aes(x = d3_5_9, y = corn_yield)` inside `geom_point()` tells R that you want to create a scatter plot where you have `d3_5_9` on the x-axis and `corn_yield` on the y-axis

]

  ]

.panel[.panel-name[Output]

.left-code[

This is what `g_fig_scatter` looks:

```{r g-fig-scatter, eval = F}  
g_fig_scatter
```
]

.right-plot[
```{r scatter-plot, ref.label = "g-fig-scatter", echo = F, out.width = "90%"}  

```
]

  ]

.panel[.panel-name[`aes()`]

.left-full[
Going back to the code,

```{r }
g_fig_scatter <- g_fig + geom_point(aes(x = d3_5_9, y = corn_yield))
```

Note that `x = d3_5_9`, `y = corn_yield` are inside `aes()`. 

<span style="color:red"> Important:</span> `aes()` is used to make the <span style='color:red'>aes</span>thetic of the figure to be a function of variables in the dataset that you told `ggplot` to use (here, `county_yield`). 

`aes(x = d3_5_9, y = corn_yield)` is telling `ggplot` to use `d3_5_9` and `corn_yield` variables in the `county_yield` dataset for the x-axis and y-axis, respectively. 

If you do not have `x = d3_5_9`, `y = corn_yield` inside `aes()`, R is going to look for `d3_5_9` and `corn_yield` themselves (but not in `county_yield`), which you have not defined.

Try

```{r eval = F}
g_fig + geom_point(x = d3_5_9, y = corn_yield)
```
]


  ]

.panel[.panel-name[summary]

.left-full[
+ `ggplot(data = dataset)` to initiate the process of creating a figure

+ add `geom_*()` to declare what kind of figure you would like to make 

+ specify what variables in the dataset to use and how they are used inside `aes()` 

+ place the `aes()` you defined above in the `geom_*()` you specified above

]

  ]

]


<!-- 
#=========================================
# Different types of figures
#=========================================
-->
  
---

class: inverse, center, middle
name: dif-geoms

# Different types of figures

<html><div style='float:left'></div><hr color='#EB811B' size=1px width=1000px></html>

---

# Different types of figures

.panelset[ 

.panel[.panel-name[Figure types]

<br>

`ggplot2` lets you create lots of different kinds of figures via various `geom_*()` functions.

+ `geom_histogram()`/`geom_density()`
+ `geom_line()`
+ `geom_boxplot()`
+ `geom_bar()`

How to specify aesthetics vary by `geom_*()`.

  ]

.panel[.panel-name[Histogram]

.left-code[
```{r eval = F}  
g_fig +
  geom_histogram(
    aes(x = corn_yield)
  )
```

`geom_histogram()` only needs `x`.
]

.right-plot[
```{r hist-ex, echo = F, out.width = "90%"}  
g_fig +
  geom_histogram(
    aes(x = corn_yield)
  )
```
]

  ]

.panel[.panel-name[Density Plot]

.left-code[
```{r density_type, eval = F}  
g_fig +
  geom_density(
    aes(x = corn_yield)
  )
```

`geom_density()` only needs `x`.
]

.right-plot[
```{r density-ex, ref.label = "density_type", echo = F, out.width = "90%"}  

```
]

  ]

.panel[.panel-name[Line plot]

.left-code[

Create a dataset first:

```{r }  
mean_yield <- county_yield %>%
  group_by(year) %>%
  summarize(
    corn_yield = mean(corn_yield, na.rm = TRUE)
  ) %>%
  filter(!is.na(year))
```

Create a line plot:

```{r line_type, eval = F}
ggplot(data = mean_yield) +
  geom_line(aes(x = year, y = corn_yield))
```

+ `geom_line()` needs `x` and `y`.

]

.right-plot[
```{r line-ex, ref.label = "line_type", echo = F, out.width = "90%"}  
```
]

  ]

.panel[.panel-name[Boxplot]

.left-code[
```{r boxplot_type, eval = F}  
ggplot(data = county_yield) +
  geom_boxplot(
    aes(x = factor(year), y = corn_yield)
  )
```

+ `geom_boxplot()` needs `x` and `y`

]

.right-plot[
```{r box-ex, ref.label = "boxplot_type", echo = F, out.width = "90%"}  

```
]
  ]

.panel[.panel-name[Bar plot]

.left-code[
```{r barplot_type, eval = F}  
ggplot(data = mean_yield) +
  geom_bar(
    aes(
      x = year,
      y = corn_yield
    ),
    stat = "identity"
  )
```

+ `geom_bar()` needs `x` and `y`

]

.right-plot[
```{r bar-ex, ref.label = "barplot_type", echo = F, out.width = "90%"}  

```
]
  ]

]

---

# Modifying how figures look

.left-full[

All the elements in the figures we have created so far are in black and white.

You can change how figure elements look by providing options inside `geom_*()`.

Here are the list of options to control the aesthetics of figures: 

+ fill
+ color
+ size
+ shape
+ linetype

Elements of figures that you can modify differ by `geom` types

The same element name can mean different things based on `geom` types

]

---

```{r fig-scatter-f, cache = F, include = F}
g_fig + #BREAK
  geom_point(
    aes(x = d3_5_9, y = corn_yield),
    color = "red", #BREAK2
    size = 0.7, #BREAK3
    shape = 0 #BREAK4
  )
```

`r chunk_reveal(
  chunk_name = "fig-scatter-f", 
  title = "# Scatter Plot",
  break_type = "non_seq", 
  display_type = "both", 
  left_assign = FALSE,
  width = c(50, 50)
  )
`

+ `color = "red"`: makes all the squares red
+ `size = 0.5`: makes the size of the squares smaller 
+ `shape = 0`: change the shape of the points (find other shapes [here](http://www.sthda.com/english/wiki/ggplot2-point-shapes))

---

```{r hist-f, cache = F, include = F}  
g_fig +
  geom_histogram(
    aes(x = corn_yield),
    color = "blue",
    fill = "green",
    size = 2,
    shape = 2
  )
```

`r chunk_reveal(
  chunk_name = "hist-f", 
  title = "# Histogram",
  #break_type = "non_seq", 
  display_type = "both", 
  left_assign = FALSE,
  width = c(50, 50)
  )
`
+ `color = "blue"`: makes all the boundary of the bars blue
+ `fill = "green"`: makes the fill of the bars green
+ `size = 2`: makes the line width of the boundary of the bars thicker 
+ `shape = 2`: does nothing
 

---


```{r box-f, cache = F, include = F}
ggplot(data = county_yield) + #BREAK
  geom_boxplot(
    aes(x = factor(year), y = corn_yield),
    color = "red", #BREAK2
    fill = "orange", #BREAK3
    size = 0.2, #BREAK4
    shape = 1 #BREAK5
  )
```

`r chunk_reveal(
  chunk_name = "box-f", 
  title = "# Box Plot",
  break_type = "non_seq", 
  display_type = "both", 
  left_assign = FALSE,
  width = c(50, 50)
  )
`

+ `color = "blue"`: makes all the boundary of the boxes red
+ `fill = "orange"`: makes the fill of the boxes orange
+ `size = 0.2`: makes the line width of the boundary of the boxes thinner 
+ `shape = 1`: does nothing

---

```{r line-f, cache = F, include = F}
ggplot(data = mean_yield) + #BREAK1
  geom_line(
    aes(x = year, y = corn_yield),
    color = "blue", #BREAK2
    size = 1.5, #BREAK3
    fill = "red", #BREAK4
    linetype = "dotted", #BREAK5
  )
```

`r chunk_reveal(
  chunk_name = "line-f", 
  title = "# Line Plot",
  break_type = "non_seq", 
  display_type = "both", 
  left_assign = FALSE,
  width = c(50, 50)
  )
`

+ `color = "blue"`: makes the line blue
+ `size = 1.5`: makes the line thicker 
+ `fill = "red"`: does nothing
+ `linetype = "dotted"`: makes the line dotted

---

# Exercises

.panelset[ 

.panel[.panel-name[Instruction]

This exercise use the `diamonds` dataset from the `ggplot2()` package. First, load the dataset and extract observations with `Premium` cut whose color is one of `E`, `I`, and `F`:

```{r diamond, cache=TRUE}
data("diamonds")
premium <- diamonds %>%
  filter(
    cut == "Premium" & color %in% c("E", "I", "F")
  )

#--- take a look ---#
premium
```

  ]

.panel[.panel-name[Exercise 1]

<br>

Using `carat` and `price` variables from `premium`, generate the figure below:

```{r diamond_fig_1, echo = FALSE, cache = TRUE, out.width = "50%"}
ggplot(data = premium) +
  geom_point(aes(x = carat, y = price), color = "red")
```

  ]

.panel[.panel-name[Exercise 2]

<br>

Using `price` variables from `premium`, generate a histogram of `price` shown below:

```{r diamond_fig_2, echo = FALSE, cache = TRUE, out.width = "50%"}
ggplot(data = premium) +
  geom_histogram(aes(x = price), fill = "white", color = "blue")
```

  ]

]


<!-- 
#=========================================
# Placing more information
#=========================================
-->
  
---

class: inverse, center, middle
name: more-info

# Placing more information in one figure

<html><div style='float:left'></div><hr color='#EB811B' size=1px width=1000px></html>

---

# Placing more information in one figure

.panelset[ 

.panel[.panel-name[Motivation]

<br>

So far, we have learned how to create popular types of figures. 

We can make a figure much more informative by making its aesthetics data-dependent.

For example, suppose you are interested in comparing the history of irrigated corn yield by state in a line plot.

So, you want to create a line for each state and make the lines distinguishable so the readers know which line is for which state like this:   

   
```{r more-info, echo = F, out.width = "60%", fig.dim = c(6, 3)}
county_yield_mean <- county_yield %>%
  group_by(state_name, year) %>%
  summarize(corn_yield = mean(corn_yield, na.rm = T))

ggplot(data = county_yield_mean) +
  geom_line(
    aes(y = corn_yield, x = year, color = state_name)
  )
```

  ]

.panel[.panel-name[How]

.left-full[

We can make the aesthetics of a figure data-dependent by specifying which variable you use for aesthetics differentiation <span style="color:red"> INSIDE </span>`aes()`.

Here is an example:

```{r how-dif, include = F}
ggplot(data = county_yield_mean) +
  geom_line(
    aes(y = corn_yield, x = year, color = state_name)
  )
```

```{r how-dif-flair, echo = F}
decorate("how-dif", eval = F) %>%
  flair("color = state_name") %>%
  knit_print.with_flair()
```

In this code, `color = state_name` is inside `aes()` and it tells R to divide the data into the groups of State and draw a line by `state_name` (by state) where the lines are color-differentiated.

A legend is automatically generated.

]

  ]

.panel[.panel-name[Let's do it]

<br>

.left-code[

Create a data set of corn yield by state-year first:

```{r }  
county_yield_mean <- county_yield %>%
  group_by(state_name, year) %>%
  summarize(corn_yield = mean(corn_yield, na.rm = T))
```

Create a plot:

```{r corn-color-dif, eval = F}
ggplot(data = county_yield_mean) +
  geom_line(
    aes(
      y = corn_yield,
      x = year,
      color = state_name #<<
    )
  )
```

]

.right-plot[
```{r do-it, ref.label = "corn-color-dif", echo = F, out.width = "90%"}  

```
]

  ]

]

---


```{r density-more, cache = F, include = F}
ggplot(data = county_yield_mean) + #BREAK
  geom_density(
    aes(
      x = corn_yield,
      fill = state_name #BREAK2
    ),
    alpha = 0.3
  )
```

`r chunk_reveal(
  chunk_name = "density-more", 
  title = "# More examples: Density Plot",
  break_type = "non_seq", 
  display_type = "both", 
  left_assign = FALSE,
  width = c(50, 50)
  )
`

---


```{r box-more, cache = F, include = F}
county_yield %>%
  filter(state_name %in% c("Nebraska", "Kansas")) %>%
  ggplot(data = .) +
  geom_boxplot(
    aes(
      x = factor(year),
      y = corn_yield,
      fill = state_name #BREAK2
    )
  )
```

`r chunk_reveal(
  chunk_name = "box-more", 
  title = "# More examples: Boxplot",
  break_type = "non_seq", 
  display_type = "both", 
  left_assign = FALSE,
  width = c(50, 50)
  )
`

---


```{r scatter-more, cache = F, include = F}
county_yield %>%
  filter(state_name %in% c("Nebraska", "Kansas")) %>%
  ggplot(data = .) + #BREAK
  geom_point(
    aes(
      x = d3_5_9,
      y = corn_yield,
      color = state_name, #BREAK2
      shape = state_name #BREAK3
    ),
    size = 0.7
  )
```

`r chunk_reveal(
  chunk_name = "scatter-more", 
  title = "# More examples: Scatter Plot",
  break_type = "non_seq", 
  display_type = "both", 
  left_assign = FALSE,
  width = c(50, 50)
  )
`

---

# Exercises

.panelset[ 


.panel[.panel-name[Exercise 1]

<br>

Using `premium`, create a scatter plot of `price` (y-axis) against `depth` (x-axis) by `clarity` as shown below:

```{r ex_2_1, echo = FALSE, cache = TRUE, out.width = "60%", fig.dim = c(6, 4)}
ggplot(data = premium) +
  geom_point(aes(y = price, x = depth, color = clarity))
```

  ]

.panel[.panel-name[Exercise 2]

<br>

Using `premium`, create density plots of `carat` by `color` as shown below (set `alpha` to 0.5):

```{r ex_2_2, echo = FALSE, cache = TRUE, out.width  =  "60%", fig.dim = c(6, 4)}
ggplot(data = premium) +
  geom_density(aes(x = carat, fill = color), alpha = 0.5)
```

  ]

]

<!-- 
#=========================================
# Faceting
#=========================================
-->
  
---

class: inverse, center, middle
name: faceting

# Faceting

<html><div style='float:left'></div><hr color='#EB811B' size=1px width=1000px></html>

---

# Faceting: Basics

.panelset[ 


.panel[.panel-name[Motivation]

Sometimes, you would like to visualize information across groups on separate panels. 

.left5[

Too much information in one panel?

```{r box_all, echo = F, cache=TRUE}
ggplot(data = county_yield) +
  geom_boxplot(
    aes(
      x = factor(year),
      y = corn_yield,
      fill = state_name
    )
  )
```
]

.right5[

On separate panels (faceting)?

```{r box-faceted-1, echo = F, cache = TRUE}  
ggplot(data = county_yield) +
  geom_boxplot(
    aes(
      x = factor(year),
      y = corn_yield
    )
  ) +
  facet_wrap(state_name ~ ., nrow = 3)
```
]

  ]

.panel[.panel-name[How]

.left-full[

We can make faceted figures by adding either `facet_wrap` or `facet_grid()` in which you specify which variable to use for faceting. 

Here is an example:

```{r how-facet, include = F}
ggplot(data = county_yield) +
  geom_boxplot(
    aes(x = factor(year), y = corn_yield)
  ) +
  facet_wrap(state_name ~ .)
```

```{r echo = F}
decorate("how-facet", eval = F) %>%
  flair("facet_wrap(state_name ~ .)") %>%
  knit_print.with_flair()
```

In this code, `facet_wrap(state_name ~ .)` is added to a simple boxplot, which tells R to make a boxplot by `state_name` (state).

What does `~ .` do?

]

  ]

]

---


```{r facet-ex, cache = F, include = F}
ggplot(data = county_yield) +
  geom_boxplot(
    aes(x = factor(year), y = corn_yield)
  ) + #BREAK
  facet_wrap(state_name ~ .)
```

`r chunk_reveal(
  chunk_name = "facet-ex", 
  title = "# Faceting: an Example",
  break_type = "user", 
  display_type = "both", 
  left_assign = FALSE,
  width = c(50, 50)
  )
`

---

# Faceting: Two-way

.panelset[ 

.panel[.panel-name[two-way faceting]

.left-full[
Two-way faceting will 

+ divide the data into groups where each group has a unique combination of the two faceting variables  

+ create a plot for each group

```{r two-facet, eval = F}
ggplot(data = county_yield) +
  geom_histogram(
    aes(x = corn_yield)
  ) +
  facet_wrap(state_name ~ year) #<<
```

This code will create a histogram of corn yield for each of the unique state-year combination.

]


 ]

.panel[.panel-name[Figure 2]

.left-code[

Filter `county_yield` to those in 2017 and 2018.

```{r }
county_yield_s <- county_yield %>%
  filter(year %in% c(2017, 2018))
```

Create a faceted density plots.

```{r two-ex, eval = F}  
ggplot(data = county_yield_s) +
  geom_histogram(
    aes(x = corn_yield)
  ) +
  facet_wrap(state_name ~ year)
```
]

.right-plot[
```{r two-ex-fig, ref.label = "two-ex", echo = F, out.width = "90%"}  

```
]

  ]

]

---


# Faceting with `facet_grid()`

.panelset[ 

.panel[.panel-name[compare]

.left5[

`facet_wrap`

```{r wrap-ex-1, fig.dim = c(4,3)}
ggplot(data = county_yield_s) +
  geom_histogram(aes(x = corn_yield)) +
  facet_wrap(state_name ~ year)
```
]

.right5[

`facet_grid`

```{r frig-ex-1, fig.dim = c(4,3)}
ggplot(data = county_yield_s) +
  geom_histogram(aes(x = corn_yield)) +
  facet_grid(state_name ~ year)
```
]


  ]

.panel[.panel-name[facet_grid()]

.left-code[

Unlike `facet_wrap()`, which side you put faceting variables matters a lot. 

+ left hand side: rows 
+ right hand side: columns 

```{r grid-ex-2, eval = F}  
ggplot(data = county_yield_s) +
  geom_histogram(aes(x = corn_yield)) +
  facet_grid(state_name ~ year)
```

In the code above, `state_name` values become the rows, and `year` values become columns.
]

.right-plot[
```{r grid-matter, ref.label = "grid-ex-2", echo = F, out.width = "90%", out.width = "90%"}  

```
]

  ]

.panel[.panel-name[order]

.left5[
```{r wrap-left, ref.label = "grid-ex-2"}

```
]

.right5[
```{r grid-right}
ggplot(data = county_yield_s) +
  geom_histogram(aes(x = corn_yield)) +
  facet_grid(year ~ state_name)
```
]


  ] <!-- panel ends here -->

.panel[.panel-name[scale]

.left-full[

`facet_grid()` allows 

+ the figures in different columns to have different scales for the x-axis (figures in the same column have the same scale for the x-axis)

+ the figures in different rows to have different scales for the y-axis (figures in the same rows have the same scale for the x-axis)

]

  ] <!-- panel ends here -->

.panel[.panel-name[free x]

.left-code[
```{r free-x, eval = F}  
ggplot(data = county_yield_s) +
  geom_histogram(aes(x = corn_yield)) +
  facet_grid(
    state_name ~ year,
    scales = "free_x" #<<
  )
```

]

.right-plot[
```{r free-x-fig, ref.label = "free-x", echo = F, out.width = "90%"}  

```
]

  ] <!-- panel ends here -->

.panel[.panel-name[free y]

.left-code[
```{r free-y, eval = F}  
ggplot(data = county_yield_s) +
  geom_histogram(aes(x = corn_yield)) +
  facet_grid(
    state_name ~ year,
    scales = "free_y" #<<
  )
```

]

.right-plot[
```{r free-y-fig, ref.label = "free-y", echo = F, out.width = "90%"}  

```
]


  ] <!-- panel ends here -->

.panel[.panel-name[both free]

.left-code[
```{r free, eval = F}  
ggplot(data = county_yield_s) +
  geom_histogram(aes(x = corn_yield)) +
  facet_grid(
    state_name ~ year,
    scales = "free" #<<
  )
```

]

.right-plot[
```{r both-free-fig, ref.label = "free", echo = F, out.width = "90%"}  

```
]

  ] <!-- panel ends here -->

.panel[.panel-name[strip label]

.left-code[

You can change strip labels using the `labeller = ` option inside `facet_grid()` (or `facet_wrap()`). 

To do this, you need to create a vector of labels you want where its element names are the corresponding values of the faceting variables.

Define labels first:

```{r }
#--- the vector values are new strip labels ---#
year_labels <- paste("Year = ", c(2017, 2018))

#--- the element names are the values to replace  ---#
names(year_labels) <- c("2017", "2018")
```

Create a faceted figure with new labels:

```{r strip-label, eval = F}  
ggplot(data = county_yield_s) +
  geom_histogram(aes(x = corn_yield)) +
  facet_grid(
    state_name ~ year,
    labeller = labeller(year = year_labels) #<<
  )
```

By `year = year_labels`, you are applying `year_labels` to the faceting variable `year`.

]

.right-plot[
```{r strip-label-fig, ref.label = "strip-label", echo = F, out.width = "90%"}  

```
]

  ] 

  <!-- panel ends here -->

.panel[.panel-name[or]


.left-code[
Or, you could just create a variable that has the values you want to use as labels and use it as a faceting variable:

```{r alt-strip-label, eval = F}  
county_yield_s %>%
  mutate(
    year_text = paste0("Year = ", year) #<<
  ) %>%
  ggplot(data = .) +
  geom_histogram(aes(x = corn_yield)) +
  facet_grid(
    state_name ~ year_text #<<
  )
```

]

.right-plot[
```{r alt-strip-label-f, ref.label = "alt-strip-label", echo = F, out.width = "90%"}  

```
]



  ] 

  <!-- panel ends here -->
]


---

# Exercises

.panelset[ 

.panel[.panel-name[Exercise 1]

<br>

Using `premium`, create scatter plots of `price` (y-axis) against `carat` (x-axis) by `color` on separate panels as shown below:

```{r ex_3_1, echo = FALSE, cache = TRUE, out.width = "50%"}
ggplot(data = premium) +
  geom_point(aes(x = carat, y = price)) +
  facet_grid(color ~ .)
```

  ]

.panel[.panel-name[Exercise 2]

<br>

Using premium, create histogram of `carat` by `color` and `clarity` on separate panels as shown below:

```{r ex_3_2, echo = FALSE, cache = TRUE, out.width = "50%"}
ggplot(data = premium) +
  geom_histogram(aes(x = carat)) +
  facet_grid(color ~ clarity)
```

  ]

]

---

# Density plot, histogram, boxplot

.panelset[ 

.panel[.panel-name[density-histogram]

Density plots and histograms convey basically the same information. 

.content-box-green[**Key difference**]:

+ Density plots are normalized version of histograms so that the area under them are 1.
+ Histograms convey the information about the number of observations in addition to the distribution 


.left5[
```{r echo = F, out.width = "80%"}
ggplot(county_yield) +
  geom_density(
    aes(x = corn_yield, fill = state_name),
    alpha = 0.4
  )
```
]

.right5[

```{r echo = F, out.width = "80%"}
ggplot(county_yield) +
  geom_histogram(
    aes(x = corn_yield, fill = state_name),
    alpha = 0.4, position = "identity"
  )
```
]

  ]

<!-- panel ends here -->

.panel[.panel-name[density-box]


.content-box-green[**Key difference**]:

+ a density-plot provides complete information about the distribution of a single variable, but important summary statistics like mean or media are not present 
+ a box in box-plot provides incomplete information about the distribution of a single variable, but it takes up much less space in a figure

.left3[

For this reason, boxplots are particularly useful when it is desirable to place the distribution information of a single variable across groups and over time in a single panel (see the figure to the right as an example).  

You can convey similar information using density plots faceted by year. But, it is often the case that full distribution information is not necessary.

]

.right7[

```{r echo = F, out.width = "80%"}
county_yield %>%
  filter(year >= 2014) %>%
  ggplot(.) +
  geom_boxplot(
    aes(y = corn_yield, x = factor(year), fill = state_name),
    alpha = 0.4
  )
```
]

  ]

<!-- panel ends here -->

]

<!-- panel set ends here -->

---

# Preparing datasets

.left-full[
 
We have seen 

+ figures where its main elements (points, lines, boxes, etc) are made color differentiated (e.g., with `aes(color = var)` inside the `geom_*()` function)
+  faceted figures  

.content-box-blue[.red[Important]: the dataset has to be in long format to create these types of figures!!]

<br>

For example consider the following dataset in a wide format:

```{r echo = F}
county_yield[, .(county_code, corn_yield, year, state_name)] %>%
  filter(year %in% c(2000, 2001)) %>%
  dcast(county_code + state_name ~ year, value.var = "corn_yield")
```

This dataset has county-level yields for Nebraska, Colorado, and Kansas stored in variables named `2000` and `2001` (they themselves represent years).

Imagine creating boxplots of corn yield fill color-differentiated by state and faceted by year.

You will have trouble with specifying `facet_grid()` because you do not have a single variable that represents `year`.

You will find that reshaping wide datasets using `pivot_longer()` is very useful in creating figures.

 ] 

<!-- 
#=========================================
# Other useful geom_* 
#=========================================
-->
  
---

class: inverse, center, middle
name: other-geoms

# Other supplementary `geom_*()`

<html><div style='float:left'></div><hr color='#EB811B' size=1px width=1000px></html>

---

# Other supplementary `geom_*()`

.panelset[ 

.panel[.panel-name[geom_*]

.left-full[

Here are the list of useful **geom_**.

+ `geom_vline()`: draw a vertical line
+ `geom_hline()`: draw a horizontal line
+ `geom_abline()`: draw a line with the specified intercept and slope
+ `geom_smooth()`: draw an OLS-estimated regression line (other regression methods available)
+ `geom_ribbon()`: create a shaded area
+ `geom_text()` and `annotate()`: add texts in the figure

We will use `g_fig_scatter` to illustrate how these functions work.

]

  ]

.panel[.panel-name[vline and hline]

.left-code[
```{r hv-line, eval = F}  
g_fig_scatter +
  geom_vline(
    xintercept = 10,
    color = "blue"
  ) +
  geom_hline(
    yintercept = 100,
    color = "red"
  )
```
]

.right-plot[
```{r hv-line-fig, ref.label = "hv-line", echo = F, out.width = "90%"}  
```
]

  ]

.panel[.panel-name[abline]

.left-code[


```{r ab-line, eval = F}  
g_fig_scatter +
  geom_abline(
    #--- a ---#
    intercept = 50,
    #--- b ---#
    slope = 4,
    color = "blue"
  )
```

$$y = a + b\times x$$

+ `intercept`: $a$
+ `slope`: $b$

]

.right-plot[
```{r ab-line-fig, ref.label = "ab-line", echo = F, out.width = "90%"}  
```
]

  ] <!-- panel ends here -->

.panel[.panel-name[smooth]

.left-code[
```{r smooth, eval = F}  
g_fig_scatter +
  geom_smooth(
    aes(
      y = corn_yield,
      x = d3_5_9
    )
  )
```

Also try

```{r , eval = F}  
g_fig_scatter +
  geom_smooth(
    aes(
      y = corn_yield,
      x = d3_5_9
    ),
    method = "lm"
  )
```

]

.right-plot[
```{r smooth-fig, ref.label = "smooth", echo = F, out.width = "90%"}  

```
]


  ] <!-- panel ends here -->

.panel[.panel-name[ribbon]

.left-code[
```{r ribbon, eval = F}  
g_fig_scatter +
  geom_ribbon(
    aes(
      x = d3_5_9,
      ymin = 100,
      ymax = 200
    ),
    fill = "green",
    alpha = 0.3
  )
```

+ `ymin`: lower bound of the ribbon
+ `ymax`: upper bound of the ribbon

Useful when drawing confidence intervals.

]

.right-plot[
```{r ribbon-fig, ref.label = "ribbon", echo = F, out.width = "90%"}  

```
]

  ] <!-- panel ends here -->

.panel[.panel-name[text]

.left-code[
```{r text-ex, eval = F}  
g_fig_scatter +
  geom_text(
    aes(
      x = d3_5_9,
      y = corn_yield,
      label = state_name,
    )
  )
```
+ `x`, `y`: position of where texts are placed
+ `label`: variable to print

]

.right-plot[
```{r text-ex-fig, ref.label = "text-ex", echo = F, out.width = "90%"}  

```
]

  ] <!-- panel ends here -->


.panel[.panel-name[annotate]

.left-code[
```{r annotate, eval = F}  
g_fig_scatter +
  annotate(
    "text",
    x = 10,
    y = 50,
    label = "Drought hurts \n a lot!!",
    size = 3,
    color = "red"
  )
```

+ `x`: where on x-axis
+ `y`: where on y-axis
+ `label`: text to print (\n break the line)
+ size: font size

]

.right-plot[
```{r annotate-fig, ref.label = "annotate", echo = F, out.width = "90%"}  

```
]

  ] <!-- panel ends here -->

]

---



# Multiple datasets in one figure

.panelset[ 

.panel[.panel-name[multiple datasets]

.left-full[

<span style="color:red">Important: </span>`data = county_yield` declared inside `ggplot()` applies to ALL the subsequent `geom_*()`s unless overwritten locally inside individual `geom_*()`s.

Try this:

```{r eval = F}
ggplot() +
  geom_point(data = county_yield, aes(y = corn_yield, x = d3_5_9)) +
  geom_smooth(aes(y = corn_yield, x = d3_5_9))
```

It is easy to use multiple datasets inside a single `ggplot` object (or a figure). 

You just need to specify what dataset to use locally inside individual `geom_*()`s.

Let's see how this works using an example of drawing the  confidence intervals around the regression lie of the following regression:

<br>

$$corn\_yield = \beta_0 + \beta_1 d3\_5\_9 + v$$

]

  ]

.panel[.panel-name[Preparation]

.left-full[

```{r }
#--- regression ---#
reg <- lm(corn_yield ~ d3_5_9, data = county_yield)

#--- find confidence interval ---#
min_d3 <- county_yield$d3_5_9 %>% min(na.rm = TRUE) # minimum d3 observed
max_d3 <- county_yield$d3_5_9 %>% max(na.rm = TRUE) # maximum d3 observed

eval_points <- data.frame(d3_5_9 = seq(min_d3, max_d3, length = 1000)) # evaluation points

ci_bound <- predict(reg, newdata = eval_points, interval = "confidence", level = 0.9) # upper and lower bound

ci_bound_data <- cbind(eval_points, ci_bound) # combine evaluation points and ci
```

```{r }
head(ci_bound_data)
```
]

  ] 

<!-- panel ends here -->

]

---


```{r mult-geom, cache = F, include = F}
ggplot() +
  #--- scatter plot ---#
  geom_point(
    data = county_yield,
    aes(y = corn_yield, x = d3_5_9)
  ) + #BREAK
  #--- regression line ---#
  geom_line(
    data = ci_bound_data,
    aes(x = d3_5_9, y = fit),
    color = "blue",
    size = 1.2
  ) + #BREAK
  #--- confidence interval ---#
  geom_ribbon(
    data = ci_bound_data,
    aes(x = d3_5_9, ymin = lwr, ymax = upr),
    fill = "red",
    alpha = 0.4
  )
```

`r chunk_reveal(
  chunk_name = "mult-geom", 
  title = "# Multiple datasets in one figure",
  break_type = "user", 
  display_type = "both", 
  left_assign = FALSE,
  width = c(50, 50)
  )
`

