---
output:
  html_document:
    code_folding: "hide"
    number_sections: true
    toc: true
    toc_depth: 2
    toc_float: true
    df_print: paged
---

# Exercise 1: Vector Creation & Subsetting

**Objective 1.1: Creation** 
- Construct a numeric vector that contains the numbers from 5 to 15.

**Required Functions:** 
- `c()`

**Solution Code 1.1:**
```{r}
numeric_vector <- c(5:15)
print(numeric_vector)
```

**Objective 1.2: Subsetting** 
- Extract the 2nd to 5th elements of the vector.

**Required Functions:** 
- `[]`

**Solution Code 1.2:**
```{r}
subset_vector <- numeric_vector[2:5]
print(subset_vector)
```

# Exercise 2: Matrix Construction & Subsetting

**Objective 2.1: Construction** 
- Build a 4x4 matrix using the numbers from 1 to 16.
- Set the column names to "W", "X", "Y", and "Z".

**Required Functions:** 
- `matrix()`, `colnames()`

**Solution Code 2.1:**
```{r}
my_matrix <- matrix(1:16, nrow = 4)
colnames(my_matrix) <- c("W", "X", "Y", "Z")
print(my_matrix)
```

**Objective 2.2: Subsetting** 
- Extract the second and third rows of the matrix.

**Required Functions:** 
- `[]`

**Solution Code 2.2:**
```{r}
subset_matrix <- my_matrix[2:3, ]
print(subset_matrix)
```

# Exercise 3: List Making & Subsetting

**Objective 3.1: Making** 
- Create a list with three elements: a numeric vector of two numbers (10, 20), a character string "apple", and a boolean value TRUE.

**Required Functions:** 
- `list()`

**Solution Code 3.1:**
```{r}
my_list <- list(numbers = c(10, 20), fruit = "apple", check = TRUE)
print(my_list)
```

**Objective 3.2: Subsetting** 
- Extract the "fruit" element from the list.

**Required Functions:** 
- `$` or `[]`

**Solution Code 3.2:**
```{r}
fruit_element <- my_list$fruit
# or fruit_element <- my_list[["fruit"]]
print(fruit_element)
```

# Exercise 4: Data Frame Crafting & Subsetting

**Objective 4.1: Crafting** 
- Produce a data frame with three columns: `ID` (11 to 15), `Name` (Anna, Ben, Cathy, Dan, Ella), and `Score` (87, 92, 78, 95, 83).

**Required Functions:** 
- `data.frame()`

**Solution Code 4.1:**
```{r}
ID <- 11:15
Name <- c("Anna", "Ben", "Cathy", "Dan", "Ella")
Score <- c(87, 92, 78, 95, 83)

my_df <- data.frame(ID, Name, Score)
print(my_df)
```

**Objective 4.2: Subsetting** 
- Extract the rows where `Score` is greater than 90.

**Required Functions:** 
- `[]`

**Solution Code 4.2:**
```{r}
high_scores <- my_df[my_df$Score > 90, ]
print(high_scores)
```

With these exercises, learners can practice both the creation and extraction (subsetting) of elements within the fundamental R data structures.