---
output:
  html_document:
    code_folding: "hide"
    number_sections: true
    toc: true
    toc_depth: 2
    toc_float: true
    df_print: paged
---

# Load Required Packages:
```{r}
library(dplyr)
library(ggplot2) # For some datasets
```

# Merging with a single key

## Exercise 1: Joining Economics Data

**Datasets:** `economics` from the `ggplot2` package and a fictitious dataset for financial events.

**Task:** Merge the `economics` dataset with a fictitious dataset that provides a financial event for specific dates. Join these datasets using `left_join` based on `date`.

**Load Dataset:**
```{r, class.source = 'fold-show'}
data(economics)

# Creating a fictitious dataset of financial events
financial_events <- tibble(
  date = as.Date(c("1970-01-01", "1980-01-01", "1990-01-01", "2000-01-01")),
  event = c("Event A", "Event B", "Event C", "Event D")
)
```

**Solution:**
```{r}
economics_with_events <- left_join(economics, financial_events, by = "date")
```

## Exercise 2: Joining Chick Weight with Feed Type

**Datasets:** `chickwts` from the `datasets` package.

**Task:** The `chickwts` dataset contains the weight of chicks alongside feed type. Create a fictitious dataset that provides pricing information for each feed type. Join these datasets based on the `feed` type.

**Load Dataset:**
```{r, class.source = 'fold-show'}
data(chickwts)

# Creating a fictitious dataset for feed pricing
feed_prices <- tibble(
  feed = unique(chickwts$feed),
  price_per_kg = runif(length(unique(chickwts$feed)), 10, 50)
)
```

**Solution:**
```{r}
chicks_with_prices <- left_join(chickwts, feed_prices, by = "feed")
```

## Exercise 3: Joining Plant Growth Data

**Datasets:** `PlantGrowth` from the `datasets` package.

**Task:** The `PlantGrowth` dataset provides information about the weight of plants under different treatment conditions. Create a fictitious dataset that assigns a scientific team responsible for each treatment type. Merge these datasets based on the `group` column.

**Load Dataset:**
```{r, class.source = 'fold-show'}
data(PlantGrowth)

# Creating a fictitious dataset for scientific teams
teams <- tibble(
  group = unique(PlantGrowth$group),
  team = c("Team Alpha", "Team Beta", "Team Gamma")
)
```

**Solution:**
```{r}
plants_with_teams <- left_join(PlantGrowth, teams, by = "group")
```

# Merging with multiple keys

## Exercise 1: Joining Students with Courses by StudentID and Semester

**Datasets:** Two fictitious datasets: one containing student enrollment details and another containing their grades.

**Task:** Join the `enrollment` dataset with the `grades` dataset using `student_id` and `semester` as the key variables.

**Load Dataset:**
```{r, class.source = "fold-show"}
enrollment <- tibble(
  student_id = c(101, 102, 103, 101),
  semester = c("Fall", "Fall", "Spring", "Spring"),
  course_code = c("ENG101", "MAT101", "SCI101", "ENG101")
)

grades <- tibble(
  student_id = c(101, 102, 103, 101),
  semester = c("Fall", "Fall", "Spring", "Spring"),
  course_code = c("ENG101", "MAT101", "SCI101", "MAT101"),
  grade = c("A", "B+", "A-", "B")
)
```

**Solution:**
```{r}
student_grades <- left_join(enrollment, grades, by = c("student_id", "semester", "course_code"))
```

## Exercise 2: Joining Employees with Project Assignments by Department and Role

**Datasets:** Two fictitious datasets: one listing employee details and another detailing their project assignments.

**Task:** Merge the `employees` dataset with the `projects` dataset using both `department` and `role` as the key variables.

**Load Dataset:**
```{r, class.source = "fold-show"}
employees <- tibble(
  emp_id = c(1, 2, 3, 4),
  name = c("Alice", "Bob", "Charlie", "David"),
  department = c("IT", "HR", "IT", "Finance"),
  role = c("Developer", "Recruiter", "Manager", "Analyst")
)

projects <- tibble(
  project_id = c("P1", "P2", "P3", "P4"),
  department = c("IT", "HR", "IT", "Finance"),
  role = c("Developer", "Recruiter", "Manager", "Manager"),
  project_name = c("App Dev", "Hiring Campaign", "IT Strategy", "Budget Planning")
)
```

**Solution:**
```{r}
employee_projects <- left_join(employees, projects, by = c("department", "role"))
```

## Exercise 3: Joining Store Transactions with Product Pricing by Date and Product Code

**Datasets:** Two fictitious datasets: one containing transaction records and another with product pricing.

**Task:** Join the `transactions` dataset with the `pricing` dataset using both `transaction_date` and `product_code` as the key variables.

**Load Dataset:**
```{r, class.source = "fold-show"}
transactions <- tibble(
  transaction_id = 1:4,
  transaction_date = as.Date(c("2023-01-01", "2023-01-02", "2023-01-01", "2023-01-03")),
  product_code = c("X001", "X002", "X001", "X003")
)

pricing <- tibble(
  price_date = as.Date(c("2023-01-01", "2023-01-02", "2023-01-01", "2023-01-03")),
  product_code = c("X001", "X002", "X003", "X003"),
  price = c(100, 150, 90, 95)
)
```

**Solution:**
```{r}
transaction_pricing <- left_join(transactions, pricing, by = c("transaction_date" = "price_date", "product_code"))
```

These exercises emphasize the use of `left_join` on multiple key variables and should offer comprehensive practice for students.