---
output:
  html_document:
    code_folding: "hide"
    number_sections: true
    toc: true
    toc_depth: 2
    toc_float: true
    df_print: paged
---

# Filter

## Exercise 1: Filtering the `mtcars` dataset

**Objective:** Filter the `mtcars` dataset for cars that have an automatic transmission (`am == 1`) and weigh more than 3,000 lbs (`wt > 3`).

**Load Dataset:**
```{r, class.source = 'fold-show'}

data(mtcars)
```

**Solution Code:**
```{r}
library(dplyr)

filtered_mtcars <- mtcars %>%
  filter(am == 1 & wt > 3)

print(filtered_mtcars)
```

## Exercise 2: Working with the `iris` dataset

**Objective:** Filter the `iris` dataset for flowers of the species `setosa` where the sepal length (`Sepal.Length`) exceeds 5 cm.

**Load Dataset:**
```{r, class.source = 'fold-show'}

data(iris)
```

**Solution Code:**
```{r}
filtered_iris <- iris %>%
  filter(Species == "setosa" & Sepal.Length > 5)

print(filtered_iris)
```

## Exercise 3: Diving into the `diamonds` dataset from the `ggplot2` package

**Objective:** Filter the dataset for diamonds with a `cut` of "Premium" and a `carat` size between 1 and 2.

**Load Dataset:**
```{r, class.source = 'fold-show'}

library(ggplot2)
data(diamonds)
```

**Solution Code:**
```{r}
filtered_diamonds <- diamonds %>%
  filter(cut == "Premium" & carat >= 1 & carat <= 2)

print(filtered_diamonds)
```

## Exercise 4: Exploring the `airquality` dataset

**Objective:** Filter the data for days in June (`Month == 6`) where the ozone level (`Ozone`) exceeded 100 (ignoring NA values).

**Load Dataset:**
```{r, class.source = 'fold-show'}

data(airquality)
```

**Solution Code:**
```{r}
filtered_airquality <- airquality %>%
  filter(Month == 6 & Ozone > 100, !is.na(Ozone))

print(filtered_airquality)
```

## Exercise 5: Investigating the `ChickWeight` dataset

**Objective:** Filter for records of chicks (`Chick`) number 1 to 5 (inclusive) and for times (`Time`) less than or equal to 10 days.

**Load Dataset:**
```{r, class.source = 'fold-show'}

data(ChickWeight)
```

**Solution Code:**
```{r}
filtered_chickweight <- ChickWeight %>%
  filter(Chick %in% c(1:5) & Time <= 10)

print(filtered_chickweight)
```

# Mutate

## Exercise 1: Modifying the `mtcars` dataset

**Objective:** Add a column named `efficiency` that calculates miles-per-gallon (`mpg`) divided by the number of cylinders (`cyl`).

**Load Dataset:**
```{r, class.source = 'fold-show'}

data(mtcars)
```

**Solution Code:**
```{r}
library(dplyr)

mutated_mtcars <- mtcars %>%
  mutate(efficiency = mpg/cyl)

print(mutated_mtcars)
```

## Exercise 2: Working with the `iris` dataset

**Objective:** Create a new column named `area` which multiplies sepal length (`Sepal.Length`) by sepal width (`Sepal.Width`).

**Load Dataset:**
```{r, class.source = 'fold-show'}

data(iris)
```

**Solution Code:**
```{r}
mutated_iris <- iris %>%
  mutate(area = Sepal.Length * Sepal.Width)

print(mutated_iris)
```

## Exercise 3: Enhancing the `diamonds` dataset from the `ggplot2` package

**Objective:** Calculate the price per carat and name the new column `price_per_carat`.

**Load Dataset:**
```{r, class.source = 'fold-show'}

library(ggplot2)
data(diamonds)
```

**Solution Code:**
```{r}
mutated_diamonds <- diamonds %>%
  mutate(price_per_carat = price/carat)

print(mutated_diamonds)
```

## Exercise 4: Adapting the `airquality` dataset

**Objective:** Convert the temperature from Fahrenheit (`Temp`) to Celsius and name the new column `TempC`. The formula is `C = (F - 32) * 5/9`.

**Load Dataset:**
```{r, class.source = 'fold-show'}

data(airquality)
```

**Solution Code:**
```{r}
mutated_airquality <- airquality %>%
  mutate(TempC = (Temp - 32) * 5/9)

print(mutated_airquality)
```

## Exercise 5: Adjusting the `ChickWeight` dataset

**Objective:** Calculate the weight change by subtracting the minimum weight from the current weight for each chick. Name the new column `weight_change`.

**Load Dataset:**
```{r, class.source = 'fold-show'}

data(ChickWeight)
```

**Solution Code:**
```{r}
mutated_chickweight <- ChickWeight %>%
  group_by(Chick) %>%
  mutate(weight_change = weight - min(weight))

print(mutated_chickweight)
```

# Group summary

## Exercise 1: Summarizing the `mtcars` dataset

**Objective:** Group by the number of cylinders (`cyl`) and compute the average miles-per-gallon (`mpg`) for each group.

**Load Dataset:**
```{r, class.source = 'fold-show'}

data(mtcars)
```

**Solution Code:**
```{r}
library(dplyr)

mtcars_grouped1 <- mtcars %>%
  group_by(cyl) %>%
  summarize(avg_mpg = mean(mpg))

print(mtcars_grouped1)
```

## Exercise 2: Working with the `iris` dataset

**Objective:** Group by flower species (`Species`) and calculate the average sepal length (`Sepal.Length`) and sepal width (`Sepal.Width`) for each species.

**Load Dataset:**
```{r, class.source = 'fold-show'}

data(iris)
```

**Solution Code:**
```{r}
iris_grouped <- iris %>%
  group_by(Species) %>%
  summarize(
    avg_sepal_length = mean(Sepal.Length),
    avg_sepal_width = mean(Sepal.Width)
  )

print(iris_grouped)
```

## Exercise 3: Analyzing the `diamonds` dataset from the `ggplot2` package

**Objective:** Group by `cut` and `color` and compute the median price for each combination.

**Load Dataset:**
```{r, class.source = 'fold-show'}

library(ggplot2)
data(diamonds)
```

**Solution Code:**
```{r}
diamonds_grouped <- diamonds %>%
  group_by(cut, color) %>%
  summarize(median_price = median(price))

print(diamonds_grouped)
```

## Exercise 4: Delving into the `airquality` dataset

**Objective:** Group by month (`Month`) and compute the maximum temperature (`Temp`) and average ozone level (`Ozone`, omitting `NA` values) for each month.

**Load Dataset:**
```{r, class.source = 'fold-show'}

data(airquality)
```

**Solution Code:**
```{r}
airquality_grouped <- airquality %>%
  group_by(Month) %>%
  summarize(
    max_temp = max(Temp),
    avg_ozone = mean(Ozone, na.rm = TRUE)
  )

print(airquality_grouped)
```

## Exercise 5: Insights from the `ChickWeight` dataset

**Objective:** Group by diet (`Diet`) and chick number (`Chick`). For each combination, compute the final weight (i.e., weight at the maximum time).

**Load Dataset:**
```{r, class.source = 'fold-show'}
data(ChickWeight)
```

**Solution Code:**
```{r}
chickweight_grouped <- ChickWeight %>%
  group_by(Diet, Chick) %>%
  summarize(final_weight = weight[which.max(Time)])

print(chickweight_grouped)
```


# Use all

### Exercise 1: Calculate Average MPG by Cylinder

**Dataset:** `mtcars` from the `datasets` package  
**Functions to use:** `filter()`, `mutate()`, `group_by()`, `summarize()`

**Task:** Filter the dataset to cars with more than 100 horsepower. Then, for these cars, calculate the average miles per gallon (`mpg`) for each number of cylinders (`cyl`).

**Load Dataset:**
```{r, class.source = 'fold-show'}
library(datasets)
data(mtcars)
```

**Solution:**
```{r}
library(dplyr)

mtcars %>%
  filter(hp > 100) %>%
  group_by(cyl) %>%
  summarize(avg_mpg = mean(mpg))
```

### Exercise 2: Adjusted Price Calculation

**Dataset:** `diamonds` from the `ggplot2` package  
**Functions to use:** `filter()`, `mutate()`, `group_by()`, `summarize()`

**Task:** Filter diamonds that are "Ideal" in `cut` and have `carat` less than 1. Calculate an adjusted price which is 90% of the original `price`. Finally, calculate the average adjusted price for each clarity level.

**Load Dataset:**
```{r, class.source = 'fold-show'}
library(ggplot2)
data(diamonds)
```

**Solution:**
```{r}
diamonds %>%
  filter(cut == "Ideal", carat < 1) %>%
  mutate(adjusted_price = price * 0.9) %>%
  group_by(clarity) %>%
  summarize(avg_adjusted_price = mean(adjusted_price))
```

### Exercise 3: Compute Average Dispersion by Gear

**Dataset:** `mtcars`  
**Functions to use:** `filter()`, `mutate()`, `group_by()`, `summarize()`

**Task:** Filter cars with 4 or 6 cylinders. Create a new column named `disp_per_cyl` that calculates the dispersion (`disp`) per cylinder (`cyl`). Then compute the average `disp_per_cyl` for each gear (`gear`) level.

**Solution:**
```{r}
mtcars %>%
  filter(cyl %in% c(4, 6)) %>%
  mutate(disp_per_cyl = disp / cyl) %>%
  group_by(gear) %>%
  summarize(avg_disp_per_cyl = mean(disp_per_cyl))
```

!