---
output:
  html_document:
    code_folding: "hide"
    number_sections: true
    toc: true
    toc_depth: 2
    toc_float: true
    df_print: paged
---

# Load Required Packages {.unnumbered}
```{r setup, message=FALSE, class.source="fold-show"}
library(lubridate)
```


# Exercise 1: Extracting Month and Year

**Dataset:** `dates_1 <- c("2023-04-05", "2019-12-31", "2022-08-14")`

**Task:** Extract the month and year from each date in the `dates_1` vector. The output should be in the format `MM-YYYY`.

**Functions to Use:** `month()`, `year()`

**Load Dataset:**

```{r, class.source="fold-show"}
dates_1 <- c("2023-04-05", "2019-12-31", "2022-08-14")
dates_1
```

**Solution:**

```{r}
formatted_dates_1 <- paste0(month(dates_1, label = TRUE), "-", year(dates_1))
formatted_dates_1
```

---

# Exercise 2: Calculating Days Difference

**Dataset:** `start_dates <- c("2021-05-01", "2019-07-19")` and `end_dates <- c("2021-05-15", "2019-08-02")`

**Task:** Calculate the number of days between the start and end dates for each pair.

**Functions to Use:** `ymd()`, `as.duration()`

**Load Dataset:**

```{r, class.source="fold-show"}
start_dates <- c("2021-05-01", "2019-07-19")
end_dates <- c("2021-05-15", "2019-08-02")
data.frame(start_dates, end_dates)
```

**Solution:**

```{r}
days_diff <- as.duration(ymd(end_dates) - ymd(start_dates))
days_diff
```

---

# Exercise 3: Rounding Up Dates

**Dataset:** `dates_3 <- c("2023-04-05 15:30:25", "2019-12-31 06:20:15")`

**Task:** Round the dates to the nearest hour.

**Functions to Use:** `round_date()`

**Load Dataset:**

```{r, class.source="fold-show"}
dates_3 <- c("2023-04-05 15:30:25", "2019-12-31 06:20:15")
dates_3
```

**Solution:**

```{r}
rounded_dates <- round_date(ymd_hms(dates_3), "hour")
rounded_dates
```

---

# Exercise 4: Checking Leap Years

**Dataset:** `years <- c(2000, 2001, 2004, 2100)`

**Task:** Check which years in the `years` vector are leap years.

**Functions to Use:** `leap_year()`

**Load Dataset:**

```{r, class.source="fold-show"}
years <- c(2000, 2001, 2004, 2100)
years
```

**Solution:**

```{r}
leap_years_check <- leap_year(years)
leap_years_check
```

---

# Exercise 5: Adding Months to Dates

**Dataset:** `dates_5 <- c("2021-11-30", "2019-02-28")`

**Task:** Add three months to each date in the `dates_5` vector.

**Functions to Use:** `ymd()`, `months()`

**Load Dataset:**

```{r, class.source="fold-show"}
dates_5 <- c("2021-11-30", "2019-02-28")
dates_5
```

**Solution:**

```{r}
new_dates <- ymd(dates_5) + months(3)
new_dates
```
