---
output:
  html_document:
    code_folding: "hide"
    number_sections: true
    toc: true
    toc_depth: 2
    toc_float: true
    df_print: paged
---

# Exercise 1: Convert from Wide to Long Format

**Dataset:** `airquality` from the `datasets` package  
**Functions to use:** `pivot_longer()`

**Task:** Convert the `airquality` dataset from wide format to long format such that the `Ozone`, `Solar.R`, `Wind`, and `Temp` columns are transformed into two columns: "variable" and "value".

**Load Dataset:**
```{r, class.source = 'fold-show'}
library(datasets)
data(airquality)
```

**Solution:**
```{r}
library(tidyr)

airquality_long <- airquality %>%
  pivot_longer(cols = c(Ozone, Solar.R, Wind, Temp), names_to = "variable", values_to = "value")
```

# Exercise 2: Convert from Long to Wide Format

**Dataset:** The long format data from the solution of Exercise 1  
**Functions to use:** `pivot_wider()`

**Task:** Convert the `airquality_long` dataset from long format back to its original wide format.

**Solution:**
```{r}
airquality_wide <- airquality_long %>%
  pivot_wider(names_from = variable, values_from = value)
```

# Exercise 3: Reshape Multiple Value Columns

**Dataset:** `USArrests` from the `datasets` package  
**Functions to use:** `pivot_longer()`

**Task:** The dataset `USArrests` contains crime data for different US states. Convert it to a long format where you have columns: "State", "Metric", and "Value". The "Metric" column should have entries like "Murder", "Assault", "UrbanPop", and "Rape".

**Load Dataset:**
```{r, class.source = 'fold-show'}
data(USArrests)
```

**Solution:**
```{r}
USArrests_long <- USArrests %>%
  pivot_longer(cols = c(Murder, Assault, UrbanPop, Rape), names_to = "Metric", values_to = "Value")
```

